---
title: "Final race analysis"
output: html_document
date: "2025-07-22"
---

```{r setup, include=FALSE}
library(tidyverse)
library(tableone)
library(gtsummary)
library(survival)
library(ggplot2)
library(readxl)
library(ggpubr)
library(splines)
library(xlsx)
library(MatchIt)
library(survminer)


setwd('/Users/to909/Desktop/Andrew')
all<- read_excel("All_edited_new_12142022.xlsx")


Baylor0913<- read_excel("Baylor0913.xlsx") %>% select(study_id,race,rrt_crrt,days_icu,cpr_given)
Indiana0913 <- read_excel("Indiana0913.xlsx") %>% select(study_id,race,rrt_crrt,days_icu,cpr_given)
Jacksonville0913<- read_excel("Jacksonville0913.xlsx") %>% select(study_id,race,rrt_crrt,days_icu,cpr_given)
Kentukey0913<- read_excel("Kentukey0913.xlsx")  %>% select(study_id,race,rrt_crrt,days_icu,cpr_given)
MCW0913<- read_excel("MCW0913.xlsx")%>% select(study_id,race,rrt_crrt,days_icu,cpr_given)
MGH0913<- read_excel("MGH0913.xlsx")%>% select(study_id,race,rrt_crrt,days_icu,cpr_given)
Michigan0913<- read_excel("Michigan0913.xlsx")%>% select(study_id,race,rrt_crrt,days_icu,cpr_given)
Oschner0913<- read_excel("Oschner0913.xlsx")%>% select(study_id,race,rrt_crrt,days_icu,cpr_given)
Rochester0913<- read_excel("Rochester0913.xlsx")%>% select(study_id,race,rrt_crrt,days_icu,cpr_given)
USC0913<- read_excel("USC0913.xlsx")%>%mutate(rrt_crrt=NA,days_icu=NA,cpr_given=NA) %>%  select(study_id,race,rrt_crrt,days_icu,cpr_given)
Yale0913<- read_excel("Yale0913.xlsx")%>% select(study_id,race,rrt_crrt,days_icu,cpr_given)
race_added <- rbind(Baylor0913,Indiana0913,Jacksonville0913,Kentukey0913,MCW0913,MGH0913,Michigan0913,Oschner0913,Rochester0913,USC0913,Yale0913) 


all <- merge(all,race_added, by="study_id",all.x=TRUE)

all <- all %>% mutate(type_of_aki=case_when(
  final_type_of_aki == 1 ~ "Prerenal",
  final_type_of_aki == 2 ~ "HRS-AKI",
  final_type_of_aki == 3 ~"ATN",
  final_type_of_aki == 4 ~"Other",
  final_type_of_aki == 5 ~"Unable to diagnosis")) %>%
  mutate(etiology_cirrhosis=case_when(
    etiology_cirrhosis == 1 ~ "Alcohol",
    etiology_cirrhosis == 2 ~ "HCV",
    etiology_cirrhosis == 3 ~ "NASH",
    etiology_cirrhosis == 5 ~ "Multifactorial",
    etiology_cirrhosis == 6 | etiology_cirrhosis == 7 | etiology_cirrhosis == 4 ~ "Other")) %>%

  mutate(
      race_cat_4 = case_when(
        ethnicity == 2 & race == 5 ~ "NHW",  # Non-Hispanic White
        ethnicity == 1 & race == 5 ~ "HW",   # Hispanic White
        ethnicity %in% c(1,2)  & race == 3 ~ "B",  # Non-Hispanic Black
        (ethnicity %in% c(1, 2) ) & (race %in% c(1, 2, 4, 6)) ~ "NWNB"
    )) %>% 

  mutate(
   
      race_cat_5 = case_when(
        ethnicity == 2 & race == 5 ~ "NHW",  # Non-Hispanic White
        ethnicity == 1 & race == 5 ~ "HW",   # Hispanic White
        ethnicity == 2 & race == 3 ~ "NHB",  # Non-Hispanic Black
        ethnicity == 1 & race == 3 ~ "HB",   # Hispanic Black
        ethnicity %in% c(1, 2) & race %in% c(1, 2, 4, 6, 7) ~ "NWNB"
    )
  
  ) %>% 
  mutate(race=case_when(
    race == 1 ~ "American Indian or Alaska Native",
    race == 2 ~ "Asian",
    race == 3 ~ "Black or African American",
    race == 4 ~ "Native Hawaiian or Other Pacific Islander",
    race == 5 ~ "White", 
    race == 6 ~ "Other", 
    race == 7 ~ "Unknown")) %>% 
  mutate(race_cat_new=case_when(
    race == "Black or African American" ~ "Black or African American",
    race == "White" ~ "White", 
    TRUE ~ "Other"
  ),
  cirrhosis=case_when(etiology_cirrhosis == 1 ~ 1,
                      is.na(etiology_cirrhosis) ~ NA_real_,
                      TRUE ~ 0),
  transplant_evaluation_outcome=case_when( is.na(Transplant_evalutation) ~ NA_real_ ,
                                           Transplant_evalutation==1 ~ 1,
                                           TRUE ~ 0
    
  ))%>% mutate(ethnicity=ifelse(ethnicity==1,"Hispanic",ifelse(ethnicity==2,"Non-Hispanic","Unknown")))

master <- all %>% filter(!race=='Unknown',!ethnicity=="Unknown")  %>% mutate(MELD_Na_baseline =replace_na(MELD_Na_baseline,median(MELD_Na_baseline,na.rm = TRUE))) %>% filter(MELD_Na_baseline>=0)



```




```{r,include=FALSE}
master <- master %>% mutate(female=ifelse(sex==2,1,0))

calculate_meld3 <- function(female, bilirubin, sodium, inr, creatinine, albumin) {
  # Return NA if any input is NA or non-positive
  if (anyNA(c(female, bilirubin, sodium, inr, creatinine, albumin)) ||
      any(c(bilirubin, sodium, inr, creatinine, albumin) <= 0)) {
    return(NA)
  }
  
  meld3 <- 1.33 * female +
           4.56 * log(bilirubin) +
           0.82 * (137 - sodium) -
           0.24 * (137 - sodium) * log(bilirubin) +
           9.09 * log(inr) +
           11.14 * log(creatinine) +
           1.85 * (3.5 - albumin) -
           1.83 * (3.5 - albumin) * log(creatinine) +
           6
  
  return(meld3)
}

master <- master %>%
  mutate(
    meld_3 = mapply(
      calculate_meld3,
      female = female,
      bilirubin = tb_admit,
      sodium = na_admit,
      inr = inr_admit,
      creatinine = creatinine_admission,
      albumin = alb_admit
    )
  )





all_var <- c(
  "age_admission", "sex", "White", "race_cat_new", "reason_for_admission",
  "loop_diuretic", "aldosterone_antagonist", "lactulose", "rifaximin",
  "prophylactic_antibiotic", "nsaids", "beta_blockers", "albumin_given_admission",
  "albumin_amount_admission", "albumin_given_prior", "diabetes", "cad", "ckd", "htn",
  "etiology_cirrhosis", "ascites_admission", "encephalopathy_admission",
  "gi_bleed_admission", "peritonitis_admission", "hcc_admission", "tips_admission",
  "lvp_admission", "alcoholic_hepatitis_admission", "na_admit", "k_admit",
  "cl_admit", "co2_admit", "bun_admit", "creatinine_admission", "urine_creatinine",
  "ca_admit", "mg_admit", "phos_admit", "peak_creatinine", "alt_admit",
  "alkphos_admit", "tb_admit", "alb_admit", "nh3_admit", "inr_admit",
  "wbc_admit", "hct_admit", "plt_admit", "urine_sodium", "fena",
  "he_grade_admission", "sbp_admission", "dbp_admission", "rrt", "rrt_hemodialysis",
  "initial_rrt", "icu_admission", "hrs_vasoconstrictor", "pressor", "intubated",
  "respiratory_failure", "liver_transplant_listed", "liver_transplant",
  "kidney_transplant", "palliative_care", "discharge_disposition", "code_status",
  "Akin", "final_type_of_aki", "MELD_baseline", "MELD_Na_baseline", "site",
  "AKI_responders", "clif_score", "MAP", "CLIF_C_Score", "aclf_grade",
  "status_90days", "time_90days", "Encephalopathy", "Alcoholic_hepatitis",
  "HCC", "New_diagnosis_of_cirrhosis", "Transplant_evalutation", "SBP",
  "aki_stage_4", "death_discharge", "baseline_creatinine", "admission_route",
  "los",
  # Added 0317
  "admission_route", "peak_creatinine", "rrt_hemodialysis", "rrt_crrt",
  "initial_rrt", "days_icu", "intubated", "death_discharge", "respiratory_failure",
  "readmissions", "liver_transplant", "kidney_transplant", "discharge_disposition",
  "cpr_given",'race_cat_4',"icu_admission", "Transplant_evalutation","liver_transplant_listed","palliative_care","code_status","type_of_aki","meld_3"
)



# all the categorical variables 
cat_var <- c(
  "sex", "White", "race_cat_new", "reason_for_admission", "loop_diuretic", "aldosterone_antagonist", "lactulose", "rifaximin",
  "prophylactic_antibiotic", "nsaids", "beta_blockers", "albumin_given_admission", "albumin_given_prior", "diabetes", "cad", 
  "ckd", "htn", "etiology_cirrhosis", "ascites_admission", "encephalopathy_admission", "gi_bleed_admission", "peritonitis_admission", 
  "hcc_admission", "tips_admission", "lvp_admission", "alcoholic_hepatitis_admission", "he_grade_admission", "rrt", 
  "rrt_hemodialysis", "initial_rrt", "icu_admission", "hrs_vasoconstrictor", "pressor", "intubated", "respiratory_failure", 
  "liver_transplant_listed", "liver_transplant", "kidney_transplant", "palliative_care", "discharge_disposition", "code_status", 
  "Akin", "final_type_of_aki", "site", "AKI_responders", "aclf_grade", "Encephalopathy", "Alcoholic_hepatitis", "HCC", 
  "New_diagnosis_of_cirrhosis", "Transplant_evalutation", "aki_stage_4", "death_discharge", "admission_route", 
  "rrt_crrt", "readmissions", "cpr_given",'race_cat_4',"type_of_aki"
)


#create numerical variables
num_var <- setdiff(all_var,cat_var)


master[, cat_var] <- lapply(master[, cat_var], as.factor)
# relevel factor #

master$race_cat_new <- as.factor(master$race_cat_new )
master$race_cat_new <- relevel(master$race_cat_new, ref= "White")

master$race_cat_4 <- relevel(master$race_cat_4, ref= "NHW")




master <- master %>% select(all_var) %>%
  filter(if_all(all_of(c(
    "age_admission", "meld_3", "etiology_cirrhosis", "hcc_admission",
    "diabetes", "cad", "ckd", "htn", "site"
  )), ~ !is.na(.)))

# check_missing <- master %>% select(all_of(c(
#     "age_admission", "meld_3", "etiology_cirrhosis", "hcc_admission",
#     "diabetes", "cad", "ckd", "htn", "site"
#   )))
# colSums(is.na(check_missing))

```

# Table 1 race_cat_new
```{r,include=FALSE}

T1_race_all <- CreateTableOne(vars = all_var[1:60],strata = "race_cat_new",includeNA = F,addOverall = TRUE,data = master, factorVars = cat_var)
#print table one 
T1_race_all_csv <- print(T1_race_all,exact = c("HCC","New_diagnosis_of_cirrhosis","kidney_transplant"), nonnormal=num_var,showAllLevels = T,missing = T,quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
#save 
write.csv(T1_race_all_csv, file = '/Users/to909/Partners HealthCare Dropbox/Tianqi Ouyang/Projects/Subanalysis_race/final/Table1_race_all_race_cat_new_1.csv')



T1_race_all <- CreateTableOne(vars = all_var[60:97],strata = "race_cat_new",includeNA = F,addOverall = TRUE,data = master, factorVars = cat_var)
#print table one 
T1_race_all_csv <- print(T1_race_all,exact = c("HCC","New_diagnosis_of_cirrhosis","kidney_transplant"), nonnormal=num_var,showAllLevels = T,missing = T,quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
#save 
write.csv(T1_race_all_csv, file = '/Users/to909/Partners HealthCare Dropbox/Tianqi Ouyang/Projects/Subanalysis_race/final/Table1_race_all_race_cat_new_2.csv')


```

# Table 1 race_cat_4
```{r,include=FALSE}
T1_race_all <- CreateTableOne(vars = all_var,strata = 'race_cat_4',includeNA = F,addOverall = TRUE,data = master %>% select(1:60,race_cat_4), factorVars = cat_var)
#print table one 
T1_race_all_csv <- print(T1_race_all, nonnormal=num_var,showAllLevels = F,missing = T,quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
#save 
write.csv(T1_race_all_csv, file = '/Users/to909/Partners HealthCare Dropbox/Tianqi Ouyang/Projects/Subanalysis_race/final/Table1_race_all_1_race_cat_4_1.csv')


T1_race_all <- CreateTableOne(vars = all_var,strata = 'race_cat_4',includeNA = F,addOverall = TRUE,data = master %>% select(60:97,race_cat_4), factorVars = cat_var)
#print table one 
T1_race_all_csv <- print(T1_race_all, nonnormal=num_var,showAllLevels = F,missing = T,quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
#save 
write.csv(T1_race_all_csv, file = '/Users/to909/Partners HealthCare Dropbox/Tianqi Ouyang/Projects/Subanalysis_race/final/Table1_race_all_1_race_cat_4_2csv')

race_table <- table(master$race_cat_4)
custom_colors <- c("#1b9e77", "#d95f02", "#7570b3", "#e7298a", "#66a61e")
pie(race_table, main = "Distribution of Race Categories", col = custom_colors)
```




# Competing risk Univariate model 1 
```{r,echo=FALSE,warning=FALSE, message=FALSE}
master$status_90days <- factor(master$status_90days)
master %>%
  select("race_cat_4","age_admission","meld_3","etiology_cirrhosis","hcc_admission","diabetes","cad","ckd","htn","site","time_90days","status_90days"
  ) %>%
  tbl_uvregression(
    method = tidycmprsk::crr,
    y = Surv(time_90days,status_90days),
    method.args=list(failcode = 1, cencode = 0),
    pvalue_fun = ~style_pvalue(.x, digits = 3),
    exponentiate = TRUE,
  ) %>%
  bold_p() %>%        # bold p-values under a given threshold (default 0.05)
  bold_labels() %>% add_n()

```

# Competing risk Multivariate model 1 
```{r,echo=FALSE,warning=FALSE, message=FALSE}
tidycmprsk::crr(
  Surv(time_90days, status_90days) ~race_cat_4 + age_admission + meld_3 +
            etiology_cirrhosis + hcc_admission + diabetes + cad + ckd + htn + site,
  failcode = 1,
  cencode = 0,
  data = master
) %>%  tbl_regression(exponentiate = TRUE,
                      pvalue_fun = ~
                        style_pvalue(.x, digits = 3)) %>% bold_p() %>%
  bold_labels() %>% add_n()
```



# Competing risk Univariate model 2 
```{r,echo=FALSE,warning=FALSE, message=FALSE}
master$status_90days <- factor(master$status_90days)
master %>%
  select("race_cat_4","age_admission","meld_3","etiology_cirrhosis","hcc_admission","site","time_90days","status_90days"
  ) %>%
  tbl_uvregression(
    method = tidycmprsk::crr,
    y = Surv(time_90days,status_90days),
    method.args=list(failcode = 1, cencode = 0),
    pvalue_fun = ~style_pvalue(.x, digits = 3),
    exponentiate = TRUE,
  ) %>%
  bold_p() %>%        # bold p-values under a given threshold (default 0.05)
  bold_labels() 

```

# Competing risk Multivariate model 2 
```{r,echo=FALSE,warning=FALSE, message=FALSE}
tidycmprsk::crr(
  Surv(time_90days, status_90days) ~race_cat_4+ age_admission + meld_3 +
            etiology_cirrhosis + hcc_admission +  site,
  failcode = 1,
  cencode = 0,
  data = master
) %>%  tbl_regression(exponentiate = TRUE,
                      pvalue_fun = ~
                        style_pvalue(.x, digits = 3)) %>% bold_p() %>%
  bold_labels() %>% add_n()
```

# KM Plot HRS cohort 
```{r,echo=FALSE,warning=FALSE, message=FALSE}
master <- master %>% mutate(death_status_90days=ifelse(status_90days==1,1,0))




master$race_cat_4 <- factor(master$race_cat_4)

fit_90days_death<- survfit(Surv(time_90days,death_status_90days) ~ race_cat_4, data = master )



result_90days_death<- ggsurvplot(
  fit_90days_death,                     # survfit object with calculated statistics.
  data = master,  # data used to fit survival curves. 
  surv.scale = "percent",
  # legend.title = "Etiology of AKI", # legend title
  # legend.labs = c("Prerenal","HRS-AKI","ATN"), # legend labels
  # palette = c("#00ABFD","#7CAE00","#F8766D"),
  pval = T,             # show p-value of log-rank test.
  pval.coord = c(75, 1),# pvalue  position 
  pval.size = 4, # pvalue size 
  conf.int = FALSE,         # show confidence intervals for 
  xlim = c(0,90),  # present narrower X axis, but not affect survival estimates.
  xlab = "Time (days)", # Xlabel.
  break.time.by = 30,     # break X axis in time intervals by 500.
  # ggtheme = theme_minimal(), # customize plot and risk table with a theme.
  tables.theme = theme_cleantable(),
  risk.table = TRUE,       # show risk table.
  risk.table.y.text.col = T, # colour risk table text annotations.
  risk.table.y.text = T, # show bars instead of names in text annotations
  risk.table.fontsize= 4,
  # in legend of risk table
)

result_90days_death$table <- result_90days_death$table + theme(axis.line = element_blank()) + ggtitle("Number of participants entering intervals")

result_90days_death

```

# HRS cohort 
```{r,include=FALSE}
master_hrs <- master %>% filter(type_of_aki=="HRS-AKI")
```



# Competing risk Univariate model 1 HRS cohort 
```{r,echo=FALSE,warning=FALSE, message=FALSE}
master_hrs$status_90days <- factor(master_hrs$status_90days)
master_hrs %>%
  select("race_cat_4","age_admission","meld_3","etiology_cirrhosis","hcc_admission","diabetes","cad","ckd","htn","site","time_90days","status_90days"
  ) %>%
  tbl_uvregression(
    method = tidycmprsk::crr,
    y = Surv(time_90days,status_90days),
    method.args=list(failcode = 1, cencode = 0),
    pvalue_fun = ~style_pvalue(.x, digits = 3),
    exponentiate = TRUE,
  ) %>%
  bold_p() %>%        # bold p-values under a given threshold (default 0.05)
  bold_labels() %>% add_n()

```

# Competing risk Multivariate model 1 HRS cohort 
```{r,echo=FALSE,warning=FALSE, message=FALSE}
tidycmprsk::crr(
  Surv(time_90days, status_90days) ~race_cat_4+ age_admission + meld_3 +
            etiology_cirrhosis + hcc_admission + diabetes + cad + ckd + htn + site,
  failcode = 1,
  cencode = 0,
  data = master_hrs
) %>%  tbl_regression(exponentiate = TRUE,
                      pvalue_fun = ~
                        style_pvalue(.x, digits = 3)) %>% bold_p() %>%
  bold_labels() %>% add_n()
```



# Competing risk Univariate model 2 HRS cohort 
```{r,echo=FALSE,warning=FALSE, message=FALSE}
master_hrs$status_90days <- factor(master_hrs$status_90days)
master_hrs %>%
  select("race_cat_4","age_admission","meld_3","etiology_cirrhosis","hcc_admission","site","time_90days","status_90days"
  ) %>%
  tbl_uvregression(
    method = tidycmprsk::crr,
    y = Surv(time_90days,status_90days),
    method.args=list(failcode = 1, cencode = 0),
    pvalue_fun = ~style_pvalue(.x, digits = 3),
    exponentiate = TRUE,
  ) %>%
  bold_p() %>%        # bold p-values under a given threshold (default 0.05)
  bold_labels() 

```

# Competing risk Multivariate model 2 HRS cohort 
```{r,echo=FALSE,warning=FALSE, message=FALSE}
tidycmprsk::crr(
  Surv(time_90days, status_90days) ~race_cat_4+ age_admission + meld_3 +
            etiology_cirrhosis + hcc_admission +  site,
  failcode = 1,
  cencode = 0,
  data = master_hrs
) %>%  tbl_regression(exponentiate = TRUE,
                      pvalue_fun = ~
                        style_pvalue(.x, digits = 3)) %>% bold_p() %>%
  bold_labels() %>% add_n()
```

# KM Plot HRS cohort 
```{r,echo=FALSE,warning=FALSE, message=FALSE}
master_hrs <- master_hrs %>% mutate(death_status_90days=ifelse(status_90days==1,1,0))




master_hrs$race_cat_4 <- factor(master_hrs$race_cat_4)

fit_90days_death<- survfit(Surv(time_90days,death_status_90days) ~ race_cat_4, data = master_hrs )



result_90days_death<- ggsurvplot(
  fit_90days_death,                     # survfit object with calculated statistics.
  data = master_hrs,  # data used to fit survival curves. 
  surv.scale = "percent",
  # legend.title = "Etiology of AKI", # legend title
  # legend.labs = c("Prerenal","HRS-AKI","ATN"), # legend labels
  # palette = c("#00ABFD","#7CAE00","#F8766D"),
  pval = T,             # show p-value of log-rank test.
  pval.coord = c(75, 1),# pvalue  position 
  pval.size = 4, # pvalue size 
  conf.int = FALSE,         # show confidence intervals for 
  xlim = c(0,90),  # present narrower X axis, but not affect survival estimates.
  xlab = "Time (days)", # Xlabel.
  break.time.by = 30,     # break X axis in time intervals by 500.
  # ggtheme = theme_minimal(), # customize plot and risk table with a theme.
  tables.theme = theme_cleantable(),
  risk.table = TRUE,       # show risk table.
  risk.table.y.text.col = T, # colour risk table text annotations.
  risk.table.y.text = T, # show bars instead of names in text annotations
  risk.table.fontsize= 4,
  # in legend of risk table
)

result_90days_death$table <- result_90days_death$table + theme(axis.line = element_blank()) + ggtitle("Number of participants entering intervals")

result_90days_death
```






