---
title: "Subanalysis race"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
library(tidyverse)
library(tableone)
library(gtsummary)
library(survival)
library(ggplot2)
library(readxl)
library(ggpubr)
library(splines)
library(xlsx)
library(MatchIt)


setwd('/Users/to909/Desktop/Andrew')
all<- read_excel("All_edited_new_12142022.xlsx")


Baylor0913<- read_excel("Baylor0913.xlsx") %>% select(study_id,race,rrt_crrt,days_icu,cpr_given)
Indiana0913 <- read_excel("Indiana0913.xlsx") %>% select(study_id,race,rrt_crrt,days_icu,cpr_given)
Jacksonville0913<- read_excel("Jacksonville0913.xlsx") %>% select(study_id,race,rrt_crrt,days_icu,cpr_given)
Kentukey0913<- read_excel("Kentukey0913.xlsx")  %>% select(study_id,race,rrt_crrt,days_icu,cpr_given)
MCW0913<- read_excel("MCW0913.xlsx")%>% select(study_id,race,rrt_crrt,days_icu,cpr_given)
MGH0913<- read_excel("MGH0913.xlsx")%>% select(study_id,race,rrt_crrt,days_icu,cpr_given)
Michigan0913<- read_excel("Michigan0913.xlsx")%>% select(study_id,race,rrt_crrt,days_icu,cpr_given)
Oschner0913<- read_excel("Oschner0913.xlsx")%>% select(study_id,race,rrt_crrt,days_icu,cpr_given)
Rochester0913<- read_excel("Rochester0913.xlsx")%>% select(study_id,race,rrt_crrt,days_icu,cpr_given)
USC0913<- read_excel("USC0913.xlsx")%>%mutate(rrt_crrt=NA,days_icu=NA,cpr_given=NA) %>%  select(study_id,race,rrt_crrt,days_icu,cpr_given)
Yale0913<- read_excel("Yale0913.xlsx")%>% select(study_id,race,rrt_crrt,days_icu,cpr_given)
race_added <- rbind(Baylor0913,Indiana0913,Jacksonville0913,Kentukey0913,MCW0913,MGH0913,Michigan0913,Oschner0913,Rochester0913,USC0913,Yale0913) 


all <- merge(all,race_added, by="study_id",all.x=TRUE)


all <- all %>% mutate(ethnicity=ifelse(ethnicity==1,"Hispanic",ifelse(ethnicity==2,"Non-Hispanic","Unknown")))
All <- all %>% mutate(type_of_aki=case_when(
  final_type_of_aki == 1 ~ "Prerenal",
  final_type_of_aki == 2 ~ "HRS-AKI",
  final_type_of_aki == 3 ~"ATN",
  final_type_of_aki == 4 ~"Other",
  final_type_of_aki == 5 ~"Unable to diagnosis")) %>%
  mutate(etiology_cirrhosis=case_when(
    etiology_cirrhosis == 1 ~ "Alcohol",
    etiology_cirrhosis == 2 ~ "HCV",
    etiology_cirrhosis == 3 ~ "NASH",
    etiology_cirrhosis == 5 ~ "Multifactorial",
    etiology_cirrhosis == 6 | etiology_cirrhosis == 7 | etiology_cirrhosis == 4 ~ "Other")) %>%
  mutate(
   
      race_cat_5 = case_when(
        ethnicity == 2 & race == 5 ~ "NHW",  # Non-Hispanic White
        ethnicity == 1 & race == 5 ~ "HW",   # Hispanic White
        ethnicity == 2 & race == 3 ~ "NHB",  # Non-Hispanic Black
        ethnicity == 1 & race == 3 ~ "HB",   # Hispanic Black
        ethnicity %in% c(1, 2) & race %in% c(1, 2, 4, 6, 7) ~ "NWNB"
    )
  
  ) %>% 
  mutate(race=case_when(
    race == 1 ~ "American Indian or Alaska Native",
    race == 2 ~ "Asian",
    race == 3 ~ "Black or African American",
    race == 4 ~ "Native Hawaiian or Other Pacific Islander",
    race == 5 ~ "White", 
    race == 6 ~ "Other", 
    race == 7 ~ "Unknown")) %>% 
  mutate(race_cat_new=case_when(
    race == "Black or African American" ~ "Black or African American",
    race == "White" ~ "White", 
    TRUE ~ "Other"
  ),
  cirrhosis=case_when(etiology_cirrhosis == 1 ~ 1,
                      is.na(etiology_cirrhosis) ~ NA_real_,
                      TRUE ~ 0),
  transplant_evaluation_outcome=case_when( is.na(Transplant_evalutation) ~ NA_real_ ,
                                           Transplant_evalutation==1 ~ 1,
                                           TRUE ~ 0
    
  ))

master <- All %>% filter(!race=='Unknown',!ethnicity=="Unknown")  %>% mutate(MELD_Na_baseline =replace_na(MELD_Na_baseline,median(MELD_Na_baseline,na.rm = TRUE))) %>% filter(MELD_Na_baseline>=0)



all_var <- c(
  "age_admission", "sex", "White", "race_cat_new", "reason_for_admission",
  "loop_diuretic", "aldosterone_antagonist", "lactulose", "rifaximin",
  "prophylactic_antibiotic", "nsaids", "beta_blockers", "albumin_given_admission",
  "albumin_amount_admission", "albumin_given_prior", "diabetes", "cad", "ckd", "htn",
  "etiology_cirrhosis", "ascites_admission", "encephalopathy_admission",
  "gi_bleed_admission", "peritonitis_admission", "hcc_admission", "tips_admission",
  "lvp_admission", "alcoholic_hepatitis_admission", "na_admit", "k_admit",
  "cl_admit", "co2_admit", "bun_admit", "creatinine_admission", "urine_creatinine",
  "ca_admit", "mg_admit", "phos_admit", "peak_creatinine", "alt_admit",
  "alkphos_admit", "tb_admit", "alb_admit", "nh3_admit", "inr_admit",
  "wbc_admit", "hct_admit", "plt_admit", "urine_sodium", "fena",
  "he_grade_admission", "sbp_admission", "dbp_admission", "rrt", "rrt_hemodialysis",
  "initial_rrt", "icu_admission", "hrs_vasoconstrictor", "pressor", "intubated",
  "respiratory_failure", "liver_transplant_listed", "liver_transplant",
  "kidney_transplant", "palliative_care", "discharge_disposition", "code_status",
  "Akin", "final_type_of_aki", "MELD_baseline", "MELD_Na_baseline", "site",
  "AKI_responders", "clif_score", "MAP", "CLIF_C_Score", "aclf_grade",
  "status_90days", "time_90days", "Encephalopathy", "Alcoholic_hepatitis",
  "HCC", "New_diagnosis_of_cirrhosis", "Transplant_evalutation", "SBP",
  "aki_stage_4", "death_discharge", "baseline_creatinine", "admission_route",
  "los",
  # Added 0317
  "admission_route", "peak_creatinine", "rrt_hemodialysis", "rrt_crrt",
  "initial_rrt", "days_icu", "intubated", "death_discharge", "respiratory_failure",
  "readmissions", "liver_transplant", "kidney_transplant", "discharge_disposition",
  "cpr_given"
)



# all the categorical variables 
cat_var <- c(
  "sex", "White", "race_cat_new", "reason_for_admission", "loop_diuretic", "aldosterone_antagonist", "lactulose", "rifaximin",
  "prophylactic_antibiotic", "nsaids", "beta_blockers", "albumin_given_admission", "albumin_given_prior", "diabetes", "cad", 
  "ckd", "htn", "etiology_cirrhosis", "ascites_admission", "encephalopathy_admission", "gi_bleed_admission", "peritonitis_admission", 
  "hcc_admission", "tips_admission", "lvp_admission", "alcoholic_hepatitis_admission", "he_grade_admission", "rrt", 
  "rrt_hemodialysis", "initial_rrt", "icu_admission", "hrs_vasoconstrictor", "pressor", "intubated", "respiratory_failure", 
  "liver_transplant_listed", "liver_transplant", "kidney_transplant", "palliative_care", "discharge_disposition", "code_status", 
  "Akin", "final_type_of_aki", "site", "AKI_responders", "aclf_grade", "Encephalopathy", "Alcoholic_hepatitis", "HCC", 
  "New_diagnosis_of_cirrhosis", "Transplant_evalutation", "SBP", "aki_stage_4", "death_discharge", "admission_route", 
  "rrt_crrt", "readmissions", "cpr_given"
)


#create numerical variables
num_var <- setdiff(all_var,cat_var)


master[, cat_var] <- lapply(master[, cat_var], as.factor)
# relevel factor #

master$race_cat_new <- as.factor(master$race_cat_new )
master$race_cat_new <- relevel(master$race_cat_new, ref= "White")




```

```{r,include=FALSE}

T1_race_all <- CreateTableOne(vars = all_var[1:60],strata = "race_cat_new",includeNA = F,addOverall = TRUE,data = master, factorVars = cat_var)
#print table one 
T1_race_all_csv <- print(T1_race_all,exact = c("HCC","New_diagnosis_of_cirrhosis","kidney_transplant"), nonnormal=num_var,showAllLevels = T,missing = T,quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
#save 
write.csv(T1_race_all_csv, file = '/Users/to909/Partners HealthCare Dropbox/Tianqi Ouyang/Projects/Subanalysis_race/Table1_race_all_1_60.csv')



T1_race_all <- CreateTableOne(vars = all_var[60:104],strata = "race_cat_new",includeNA = F,addOverall = TRUE,data = master, factorVars = cat_var)
#print table one 
T1_race_all_csv <- print(T1_race_all,exact = c("HCC","New_diagnosis_of_cirrhosis","kidney_transplant"), nonnormal=num_var,showAllLevels = T,missing = T,quote = FALSE, noSpaces = TRUE, printToggle = FALSE)
#save 
write.csv(T1_race_all_csv, file = '/Users/to909/Partners HealthCare Dropbox/Tianqi Ouyang/Projects/Subanalysis_race/Table1_race_all_60_104.csv')


```




# Fine and Gray  event: death competing risk: liver transplant  multivariate 
```{r,echo=FALSE,warning=FALSE, message=FALSE,include=T}

master$status_90days <- as.factor(master$status_90days)
tidycmprsk::crr(Surv(time_90days,status_90days) ~race_cat_new+age_admission+sex+MELD_Na_baseline+alb_admit+baseline_creatinine+encephalopathy_admission+alcoholic_hepatitis_admission,failcode=1,cencode=0, data = master) %>% tbl_regression(exp = TRUE)%>%
  bold_p(0.01) %>%        # bold p-values under a given threshold (default 0.05)
  bold_labels() %>% add_n()
```


# Fine and Gray  event: death competing risk: liver transplant  multivariate 
```{r,echo=FALSE,warning=FALSE, message=FALSE,include=T}

master$status_90days <- as.factor(master$status_90days)
tidycmprsk::crr(Surv(time_90days,status_90days) ~hispanic_race+age_admission+sex+MELD_Na_baseline+alb_admit+baseline_creatinine+encephalopathy_admission+alcoholic_hepatitis_admission,failcode=1,cencode=0, data = master) %>% tbl_regression(exp = TRUE)%>%
  bold_p(0.01) %>%        # bold p-values under a given threshold (default 0.05)
  bold_labels() %>% add_n()
```



## multivariate logistic regression outcome icu_admission
```{r,echo=FALSE,warning=FALSE, message=FALSE,include=T}
glm(icu_admission ~ hispanic_race+age_admission+sex+MELD_Na_baseline+alb_admit+encephalopathy_admission+alcoholic_hepatitis_admission+htn+cad+ckd, family = binomial(link = 'logit'),data=master)  %>%  tbl_regression(exponentiate = TRUE,
                       pvalue_fun = ~style_pvalue(.x, digits = 3)) %>% bold_p() %>% # bold p-values under a given threshold (default 0.05)
  bold_labels()%>% 
  modify_spanning_header(
    c(estimate, ci, p.value) ~  
      "**icu_admission mutilvariate logistic regression**")%>% add_n()

```

## multivariate logistic regression outcome transplant_evaluation_outcome
```{r,echo=FALSE,warning=FALSE, message=FALSE,include=T}
glm(transplant_evaluation_outcome ~ hispanic_race+age_admission+sex+MELD_Na_baseline+alb_admit+encephalopathy_admission+alcoholic_hepatitis_admission+htn+cad+ckd, family = binomial(link = 'logit'),data=master)  %>%  tbl_regression(exponentiate = TRUE,
                       pvalue_fun = ~style_pvalue(.x, digits = 3)) %>% bold_p() %>% # bold p-values under a given threshold (default 0.05)
  bold_labels()%>% 
  modify_spanning_header(
    c(estimate, ci, p.value) ~  
      "**transplant_evaluation_outcome mutilvariate logistic regression**") %>% add_n()

```

## multivariate logistic regression outcome transplant_evaluation_outcome
```{r,echo=FALSE,warning=FALSE, message=FALSE,include=T}
glm(transplant_evaluation_outcome ~ hispanic_race+age_admission+sex+MELD_Na_baseline+alb_admit+encephalopathy_admission+alcoholic_hepatitis_admission+htn+cad+ckd, family = binomial(link = 'logit'),data=master)  %>%  tbl_regression(exponentiate = TRUE,
                       pvalue_fun = ~style_pvalue(.x, digits = 3)) %>% bold_p() %>% # bold p-values under a given threshold (default 0.05)
  bold_labels()%>% 
  modify_spanning_header(
    c(estimate, ci, p.value) ~  
      "**transplant_evaluation_outcome mutilvariate logistic regression**")%>% add_n()

```


## multivariate logistic regression outcome rrt
```{r,echo=FALSE,warning=FALSE, message=FALSE,include=T}

glm(rrt ~ hispanic_race+age_admission+sex+MELD_Na_baseline+alb_admit+encephalopathy_admission+alcoholic_hepatitis_admission+htn+cad+ckd, family = binomial(link = 'logit'),data=master)  %>%  tbl_regression(exponentiate = TRUE,
                       pvalue_fun = ~style_pvalue(.x, digits = 3)) %>% bold_p() %>% # bold p-values under a given threshold (default 0.05)
  bold_labels()%>% 
  modify_spanning_header(
    c(estimate, ci, p.value) ~  
      "**rrt mutilvariate logistic regression**")%>% add_n()

```



############### race

## multivariate logistic regression outcome icu_admission
```{r,echo=FALSE,warning=FALSE, message=FALSE,include=T}
glm(icu_admission ~ race_cat_new+age_admission+sex+MELD_Na_baseline+alb_admit+encephalopathy_admission+alcoholic_hepatitis_admission+htn+cad+ckd, family = binomial(link = 'logit'),data=master)  %>%  tbl_regression(exponentiate = TRUE,
                       pvalue_fun = ~style_pvalue(.x, digits = 3)) %>% bold_p() %>% # bold p-values under a given threshold (default 0.05)
  bold_labels()%>% 
  modify_spanning_header(
    c(estimate, ci, p.value) ~  
      "**icu_admission mutilvariate logistic regression**")%>% add_n()

```

## multivariate logistic regression outcome transplant_evaluation_outcome
```{r,echo=FALSE,warning=FALSE, message=FALSE,include=T}
glm(transplant_evaluation_outcome ~ race_cat_new+age_admission+sex+MELD_Na_baseline+alb_admit+encephalopathy_admission+alcoholic_hepatitis_admission+htn+cad+ckd, family = binomial(link = 'logit'),data=master)  %>%  tbl_regression(exponentiate = TRUE,
                       pvalue_fun = ~style_pvalue(.x, digits = 3)) %>% bold_p() %>% # bold p-values under a given threshold (default 0.05)
  bold_labels()%>% 
  modify_spanning_header(
    c(estimate, ci, p.value) ~  
      "**transplant_evaluation_outcome mutilvariate logistic regression**") %>% add_n()

```

## multivariate logistic regression outcome transplant_evaluation_outcome
```{r,echo=FALSE,warning=FALSE, message=FALSE,include=T}
glm(transplant_evaluation_outcome ~ race_cat_new+age_admission+sex+MELD_Na_baseline+alb_admit+encephalopathy_admission+alcoholic_hepatitis_admission+htn+cad+ckd, family = binomial(link = 'logit'),data=master)  %>%  tbl_regression(exponentiate = TRUE,
                       pvalue_fun = ~style_pvalue(.x, digits = 3)) %>% bold_p() %>% # bold p-values under a given threshold (default 0.05)
  bold_labels()%>% 
  modify_spanning_header(
    c(estimate, ci, p.value) ~  
      "**transplant_evaluation_outcome mutilvariate logistic regression**")%>% add_n()

```


## multivariate logistic regression outcome rrt
```{r,echo=FALSE,warning=FALSE, message=FALSE,include=T}
table(master$rrt,master$aki_stage_4)

glm(rrt ~ race_cat_new+age_admission+sex+MELD_Na_baseline+alb_admit+encephalopathy_admission+alcoholic_hepatitis_admission+htn+cad+ckd, family = binomial(link = 'logit'),data=master)  %>%  tbl_regression(exponentiate = TRUE,
                       pvalue_fun = ~style_pvalue(.x, digits = 3)) %>% bold_p() %>% # bold p-values under a given threshold (default 0.05)
  bold_labels()%>% 
  modify_spanning_header(
    c(estimate, ci, p.value) ~  
      "**rrt mutilvariate logistic regression**")%>% add_n()

```



